---
title: "SBARC AIS Cargo Summaries 2"
output:
  html_notebook: default
  html_document: default
---

Some summary data for cargo ships passing through the Santa Barbara Channel from January - March 2018. A single transit or trip consists of the vessel travelling the length of the channel in a designated shipping lane **95%** of the time. 

```{r setup, include=FALSE, message=FALSE}
#load libraries
library(pacman)
p_load(dplyr)
p_load(ggplot2)
p_load(lubridate)
p_load(cowplot)
p_load(gridExtra)
p_load(RColorBrewer)
p_load(scales)
library(knitr)
library(kableExtra)

knitr::opts_chunk$set(echo = FALSE)
```

###Number of Individual Ships
```{r echo=FALSE, message=FALSE}
cargo_tanker_summary <- readRDS("cargo_tankers_lane_transits_summary.RDS")

cargo_tanker_summary$type_general <- "Cargo"
cargo_tanker_summary$type_general[cargo_tanker_summary$ship_type>=80]<- "Tanker"

cargo_tankers_num <- cargo_tanker_summary%>%
  ungroup()%>%
  group_by(type_general)%>%
  summarize(number_vessels=length(unique(name)), num_trips=length(unique(mmsi_trip)), avg_time=mean(time_diff), max_time=max(time_diff), min_time=min(time_diff))%>%
  select("Vessel type"= type_general, "Number of Vessels"=number_vessels, "Number of Transits"=num_trips, "Average Transit Time (hrs)"=avg_time, "Max Transit Time (hrs)"=max_time, "Min Transit Time"=min_time)

kable(cargo_tankers_num, "html") %>%
  kable_styling(full_width = F)

```

##Number of monthly cargo transits
```{r echo=FALSE, message=FALSE}
#get just the time
cargo_tanker_summary$min_hour <- as.POSIXct(format(as.POSIXlt(cargo_tanker_summary$min_time),format = '%T'), format="%H:%M:%S")

#create day/night column
cargo_tanker_summary$dayNight <- ifelse(cargo_tanker_summary$min_hour > as.POSIXct('2018-05-03 04:00:00 PDT') & cargo_tanker_summary$min_hour < as.POSIXct('2018-05-03 16:00:00 PDT'), 'Day', 'Night')
cargo_tanker_summary$month <- month(cargo_tanker_summary$max_time, label = TRUE, abbr=FALSE) 

cargo_only_summary <- subset(cargo_tanker_summary, type_general=="Cargo")

cargo_only_monthly <- cargo_only_summary %>%
  ungroup()%>%
  group_by(month)%>%
  summarize(number_vessels=length(unique(name)), num_trips=length(unique(mmsi_trip)))%>%
  select("Month"=month, "Number of Vessels"= number_vessels, "Number of Transits"=num_trips)

kable(cargo_only_monthly, "html") %>%
  kable_styling(full_width = F)
```



###Number of cargo transits
```{r echo=FALSE, message=FALSE}

ggplot(data=cargo_only_summary ) +
  geom_bar(mapping=aes(x=month, fill=dayNight), position="dodge")+
  xlab("")+
  ylab("Number of Trips") +
  guides(fill=guide_legend(title="Trip time of day"))+scale_fill_manual(values=c("#ABDDA4", "#3288BD"))
  

```

##Top 8 Shipping Companies
```{r echo=FALSE, message=FALSE}


shipping_company_summary <- cargo_tanker_summary%>%
  ungroup()%>%
  filter(type_general=="Cargo")%>%
  group_by(company_x)%>%
  summarize(num_vessels=length(unique(mmsi)), num_transits=length(unique(mmsi_trip)))%>%
  ungroup()%>%
  mutate(fraction_vessels=(num_vessels/sum(num_vessels)), fraction_transits=(num_transits/sum(num_transits)))%>%
  arrange(desc(fraction_transits))%>%
  mutate(num=seq(1:n()))%>%
  mutate(top_8=as.character(company_x))


shipping_company_summary$top_8[shipping_company_summary$num > 8] <- "Other"


#create df with top 8 + other label matchign with mmsi
top_8_categories <- cargo_tanker_summary %>%
  ungroup()%>%
  filter(ship_type< 80)%>%
  group_by(mmsi, company_x)%>%
  summarize(num_transits=length(unique(mmsi_trip)))%>%
  left_join(shipping_company_summary, by="company_x")%>%
  select(mmsi,top_8)

#create a table of the top 8 companies
shipping_company_top_8 <- shipping_company_summary %>%
  top_n(8, fraction_transits)%>%
  mutate(fraction_vessels=percent(fraction_vessels), fraction_transits=percent(fraction_transits))
  #select("Company"=company_x, "Number of Vessels" = num_vessels, "Number of Tranists"= num_transits, "Fraction of Total Vessels"= fraction_vessels, "Fraction of Total Transits" = fraction_transits)

 shipping_company_top_8 %>%
   mutate(
     num_vessels=cell_spec(num_vessels, align="center"),
     num_transits=cell_spec(num_transits, align="center"),
     fraction_vessels = cell_spec(fraction_vessels, "html", align="r"), 
     fraction_transits = cell_spec(fraction_transits, "html", align="r"))%>%
   #mutate(blank=" ")%>%
   select("Company"=company_x, "Number of Vessels" = num_vessels, "Number of Tranists"= num_transits, "Fraction of Total Vessels"= fraction_vessels, "Fraction of Total Transits" = fraction_transits)%>%
   kable("html", escape=F) %>%
   kable_styling(full_width = F)%>%
   column_spec(1, width="10em")%>%
   column_spec(2, width="8em")%>%
    column_spec(3, width="8em")%>%
   column_spec(4, width="8em")%>%
   column_spec(5, width="8em")
```



```{r echo=FALSE, message=FALSE}

num_transits <- shipping_company_summary %>%
  group_by(top_8)%>%
  summarize(ttl_transits=sum(num_transits))%>%
  mutate(perc_tranist=ttl_transits/sum(ttl_transits))%>%
  arrange(ttl_transits)

#reorder based on number of total transits
num_transits <- transform(num_transits, top_8=reorder(top_8, -ttl_transits) ) 

bp <- ggplot(num_transits, aes(x=2, y=ttl_transits, fill=top_8))+geom_bar(width=1, stat="identity")+
  xlim(0.5, 2.5)

pie <- bp +coord_polar(theta="y")
pie + scale_fill_brewer(palette="Spectral")+
coord_polar(theta = "y")+
labs(x=NULL, y=NULL)+
labs(fill="") +
ggtitle("Number of Transits by Shipping Company")+
theme_bw()+
theme(plot.title = element_text(face="bold",family=c("sans"),size=15),
legend.text=element_text(size=10),
axis.ticks=element_blank(),
axis.text=element_blank(),
axis.title=element_blank(),
panel.grid=element_blank(),
panel.border=element_blank())


```



###Plot of All AIS points > 16 knots and < 16 knots for each shipping company


```{r echo=FALSE, message=FALSE}
#This is an absolute mess, I need to figure it out how to put multiple donut plots in a frame - try facet and grid.arrange???

#get individual AIS points
cargo_tanker_lane_transits_refined_30 <- readRDS("cargo_tanker_lane_transits_refined_30.RDS")
cargo_tanker_lane_transits_refined_30$mmsi <- as.character(cargo_tanker_lane_transits_refined_30$mmsi)

xx <- cargo_tanker_lane_transits_refined_30 %>%
  filter(ship_type < 80)%>%
  left_join(top_8_categories, by="mmsi") 

xx$speed_category <- "Greater than 16 knots"
xx$speed_category[xx$speed <= 16] <- "Less than 16 knots"

xx_group <- xx %>%
  group_by(top_8, speed_category)%>%
  summarize(number_points=n())%>%
  filter(top_8 != "Other")

#should do all of below in a loop instead
xx_group_1<- xx_group %>%
  filter(top_8=="Company 1")

xx_group_2<- xx_group %>%
  filter(top_8=="Company 2")

xx_group_3<- xx_group %>%
  filter(top_8=="Company 3")

xx_group_4<- xx_group %>%
  filter(top_8=="Company 4")

xx_group_5<- xx_group %>%
  filter(top_8=="Company 5")

xx_group_6<- xx_group %>%
  filter(top_8=="Company 6")

xx_group_7<- xx_group %>%
  filter(top_8=="Company 7")

xx_group_8<- xx_group %>%
  filter(top_8=="Company 8")

#brewer.pal(n = 9, name = "Spectral")

bp1 <- ggplot(xx_group_1, aes(x=2, y=number_points, fill=speed_category))+geom_bar(width=1, stat="identity")+
  xlim(0.5, 2.5) +coord_polar(theta="y")+ 
  coord_polar(theta = "y")+
  scale_fill_manual(values = c("red", "blue"))+
labs(x=NULL, y=NULL)+
labs(fill="") +
ggtitle("Company 1 (146 transits)")+
theme_bw()+
theme(plot.title = element_text(face="bold",family=c("sans"),size=15),
legend.text=element_text(size=10),
#legend.position = "bottom",
axis.ticks=element_blank(),
axis.text=element_blank(),
axis.title=element_blank(),
panel.grid=element_blank(),
panel.border=element_blank())+
  scale_fill_manual(values=c("#D53E4F", "#3288BD"))

bp2 <- ggplot(xx_group_2, aes(x=2, y=number_points, fill=speed_category))+geom_bar(width=1, stat="identity")+
  xlim(0.5, 2.5) +coord_polar(theta="y")+ 
  coord_polar(theta = "y")+
  scale_fill_manual(values = c("red", "blue"))+
labs(x=NULL, y=NULL)+
labs(fill="") +
ggtitle("Company 2 (65 transits)")+
theme_bw()+
theme(plot.title = element_text(face="bold",family=c("sans"),size=15),
legend.text=element_text(size=10),
#legend.position = "bottom",
axis.ticks=element_blank(),
axis.text=element_blank(),
axis.title=element_blank(),
panel.grid=element_blank(),
panel.border=element_blank())+
  scale_fill_manual(values=c("#D53E4F", "#3288BD"))

bp3 <- ggplot(xx_group_3, aes(x=2, y=number_points, fill=speed_category))+geom_bar(width=1, stat="identity")+
  xlim(0.5, 2.5) +coord_polar(theta="y")+ 
  coord_polar(theta = "y")+
  scale_fill_manual(values = c("red", "blue"))+
labs(x=NULL, y=NULL)+
labs(fill="") +
ggtitle("Company 3 (64 transits)")+
theme_bw()+
theme(plot.title = element_text(face="bold",family=c("sans"),size=15),
legend.text=element_text(size=10),
#legend.position = "bottom",
axis.ticks=element_blank(),
axis.text=element_blank(),
axis.title=element_blank(),
panel.grid=element_blank(),
panel.border=element_blank())+
  scale_fill_manual(values=c("#D53E4F", "#3288BD"))

bp4 <- ggplot(xx_group_4, aes(x=2, y=number_points, fill=speed_category))+geom_bar(width=1, stat="identity")+
  xlim(0.5, 2.5) +coord_polar(theta="y")+ 
  coord_polar(theta = "y")+
  scale_fill_manual(values = c("red", "blue"))+
labs(x=NULL, y=NULL)+
labs(fill="") +
ggtitle("Company 4 (28 transits)")+
theme_bw()+
theme(plot.title = element_text(face="bold",family=c("sans"),size=15),
legend.text=element_text(size=10),
#legend.position = "bottom",
axis.ticks=element_blank(),
axis.text=element_blank(),
axis.title=element_blank(),
panel.grid=element_blank(),
panel.border=element_blank())+
  scale_fill_manual(values=c("#D53E4F", "#3288BD"))

bp5 <- ggplot(xx_group_5, aes(x=2, y=number_points, fill=speed_category))+geom_bar(width=1, stat="identity")+
  xlim(0.5, 2.5) +coord_polar(theta="y")+ 
  coord_polar(theta = "y")+
  scale_fill_manual(values = c("red", "blue"))+
labs(x=NULL, y=NULL)+
labs(fill="") +
ggtitle("Company 5 (26 transits)")+
theme_bw()+
theme(plot.title = element_text(face="bold",family=c("sans"),size=15),
legend.text=element_text(size=10),
#legend.position = "bottom",
axis.ticks=element_blank(),
axis.text=element_blank(),
axis.title=element_blank(),
panel.grid=element_blank(),
panel.border=element_blank())+
  scale_fill_manual(values=c("#D53E4F", "#3288BD"))

bp6 <- ggplot(xx_group_6, aes(x=2, y=number_points, fill=speed_category))+geom_bar(width=1, stat="identity")+
  xlim(0.5, 2.5) +coord_polar(theta="y")+ 
  coord_polar(theta = "y")+
  scale_fill_manual(values = c("red", "blue"))+
labs(x=NULL, y=NULL)+
labs(fill="") +
ggtitle("Company 6 (26 transits)")+
theme_bw()+
theme(plot.title = element_text(face="bold",family=c("sans"),size=15),
legend.text=element_text(size=10),
#legend.position = "bottom",
axis.ticks=element_blank(),
axis.text=element_blank(),
axis.title=element_blank(),
panel.grid=element_blank(),
panel.border=element_blank())+
  scale_fill_manual(values=c("#D53E4F", "#3288BD"))

bp7 <- ggplot(xx_group_7, aes(x=2, y=number_points, fill=speed_category))+geom_bar(width=1, stat="identity")+
  xlim(0.5, 2.5) +coord_polar(theta="y")+ 
  coord_polar(theta = "y")+
  scale_fill_manual(values = c("red", "blue"))+
labs(x=NULL, y=NULL)+
labs(fill="") +
ggtitle("Company 7 (24 transits)")+
theme_bw()+
theme(plot.title = element_text(face="bold",family=c("sans"),size=15),
legend.text=element_text(size=10),
#legend.position = "bottom",
axis.ticks=element_blank(),
axis.text=element_blank(),
axis.title=element_blank(),
panel.grid=element_blank(),
panel.border=element_blank())+
  scale_fill_manual(values=c("#D53E4F", "#3288BD"))

bp8 <- ggplot(xx_group_8, aes(x=2, y=number_points, fill=speed_category))+geom_bar(width=1, stat="identity")+
  xlim(0.5, 2.5) +coord_polar(theta="y")+ 
  coord_polar(theta = "y")+
  scale_fill_manual(values = c("red", "blue"))+
labs(x=NULL, y=NULL)+
labs(fill="") +
ggtitle("Company 8 (22 transits)")+
theme_bw()+
theme(plot.title = element_text(face="bold",family=c("sans"),size=15),
legend.text=element_text(size=10),
#legend.position = "bottom",
axis.ticks=element_blank(),
axis.text=element_blank(),
axis.title=element_blank(),
panel.grid=element_blank(),
panel.border=element_blank())+
  scale_fill_manual(values=c("#D53E4F", "#3288BD"))

bp1
bp2
bp3
bp4
bp5
bp6
bp7
bp8


```